Left-Behind Children's Educational Performance Drops Slightly After Parents Migrated

### 2022-06-25 MScThesisText.Rmd
write up main text, put more visuals in appendix

---
title             : "Newly Left-Behind Children's Educational Performance Drops Slightly After Parents Migrated"
shorttitle        : "Left-Behind Children"

@@

Millions of children are resided in their home communities but separated from their parents as either one or both parents have migrated for work. These are the so-called "left-behind children" or "stay-behind children." They are common in migrant-sending regions, such as Latin America, Sub-Sahara Africa, East Europe, and large parts of Asia. Researchers estimated that one-sixth of Mexican children [@dewaard2018], one-fifth of Bulgarian children [@popova2018], and one-fourth of Chinese children were living apart from their migrating parents [@nbs2017].

How are stay-behind children faring when parents have migrated? Scholars have obtained mixed results due to the differential effects of parental migration. Consider children's education. According to the review articles by @debrauw2019 and @vanhook2020, labor migrants send remittances back, which can fund children's schooling as well as improve their nutritional and living conditions. The inflow of remittances, however, cannot substitute for the loss of parental care. Parental migration decreases the amount and quality of parental guidance, protection, and support. It is uncertain how these two opposing effects together shape children's educational performance, which merits empirical investigation.

A challenge arises when we want to assess the causality between parent's migration and children's educational achievement. Experiments can unveil causal relations, but it is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Alternatively, we can infer causality from a quasi-experimental design -- that is, when the cause is on par with random assignment to participants in the analysis, conditional on identification assumptions [@cunningham2021].

In this thesis, my research question is: To what extent do parental migration affect the educational performance of left-behind children in the short term? I chose China as the case and analyzed data from a two-wave panel survey. The nationally representative sample includes children in both urban and rural areas. I developed a difference-in-differences model for causal inference and applied propensity score matching for robustness check. The outcome was measured with children's cognitive test score and academic exam scores. I demonstrate that parental migration reduces the educational performance of children staying behind, but only to a small extent.

@@

The other theoretical model is the family disruption model. Children flourish under parental nurturing care. This term refers to the "stable environments that promote health and adequate nutrition, protect from threats, and provide opportunities for learning and responsive, emotionally supportive and developmentally enriching relationships" [@black2021, p.1]. Parental migration decreases nurturing care, thereby hindering children's developmental outcomes. The loss of nurturing care slows children's accumulation of human capital. Hong talked to a group of left-behind children at a rural school [@hong2019]. These students expressed the utter sense of loneliness "with no advocate, no support and with no one looking out for them" (p. 13). Many of them felt indifferent to school grades, and some even longed to ditch school to pursue a free, independent life (p. 14).

Researchers obtained mixed results regarding how parental migration affects children's education. @botezat2020 studied the case of Romania and found a positive effect of parents' migration on academic grades of adolescents aged eleven to fifteen. @wang2019 collected data from junior high school students in central China and found no effect of parental on left-behind children's achievement in mathematical test. Studying four Chinese provinces, @debrauw2017 showed that as ample migration opportunities appeared in villages, enrollment to senior high school decreased. @nguyen2016 demonstrated that parental migration was associated with lower cognitive capabilities among Indian and Vietnamese children aged five to eight. The inconsistency can result from factors such as the left-behind children's age and gender, the migrant parent's original domicile and eventual destination, and the arrangement of which parent migrate for how long.

@@

Researchers have studied whether the impact differ between mother absence and father absence in the context of labor migration. @xu2019 adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had "little or no association with negative outcomes" on children's academic and cognitive performances. Only-mother absence, however, showed a "strong association with negative outcomes" (p. 1646). @chen2019 conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that mother migration was "negatively associated with children's social competence and academic performance," while father migration had no direct effects on the two outcomes (pp. 860-861).

In China, the social norms designate men as breadwinners and women as caregivers. Studies noted that in some Asian and Latin American societies, children are more likely to accept fathers' migration than mothers' [@murphy2022]. Left-behind children may even resent their migrant mothers, viewing them as transgressing on the expected care-giving roles. I propose the second hypothesis:

$H_2$ In the short term, mother migration negatively affect the cognitive development and academic performance of children left-behind, while father migration does not show a effect.

@@

Three features make China a suitable case for studying left-behind children. The first feature is the huge scale of internal labor migration. According to the National Bureau of Statistics, the year 2021 recorded 292.5 million in-country migrant workers [@nbs2022]. They form the backbone of essential industries, such as retail, delivery, transportation, and manufacturing. The number of left-behind Chinese children reached 68.7 million in 2015 [@nbs2017]. The second feature is the neoliberal welfare regime. Neoliberalism extends the logic of the market to most spheres of human life. Driven by the neoliberal doctrine, the government shifts its responsibilities of social services and welfare to individuals and communities. Since China initiated pro-market reforms in the 1980s, the state has reduced its welfare provisions and passed them to individuals and families [@duckett2020]. The third is the authoritarian control against citizens. According to @gu2021 and @zhang2018, the hukou (household registration) system controls people's mobility by binding a household's public welfare access with its place of registry. One can move out of town freely but cannot change hukou freely. For rural household members moving to the city, it takes months -- if not years -- to apply for a new hukou. Before receiving urban hukou, they only enjoy limited access to public service in the host city, including public education for children. Migrant laborers had organized strikes and protests, only to meet with police crackdown [@chan2017]. Facing financial and policy constraints, many migrant laborers have to leave children in their hometowns [@gu2021; @zhang2018].

The existing research about China's left-behind children has three major shortcomings. The first is that most literature focuses on children left behind in rural areas while neglecting the urban counterparts. The National Bureau of Statistics stated that the number of urban left-behind children had been growing and expected that the trend would continue [@nbs2017]. My sample includes urban left-behind children as well. The second shortcoming lies in sampling and research design. Some studies suffer from a limited sample size of less than one thousand participants or a limited geographic coverage of one to a few towns [@liu2021; @yang2022; @shu2021; @jin2020]. Others only estimate correlation using cross-sectional data [@akezhuoli2022; @jin2020]. These studies could have applied causal inference techniques like fixed effect, matching, or inverse probability of treatment weighting. Others s The third shortcoming lies in the measurement of outcome. @chang2019 and @wang2019 measured students' educational performance with the test score of only one academic subject, namely mathematics. To tackle these limitations, I adopted a nationally representative two-wave panel dataset. I included the test results of multiple subjects to measure comprehensive academic achievement.

@@

I explored two categories of educational performance: cognitive and academic abilities. One key dependent variable, the student's cognitive skill, was measured by a fifteen-minute standardized test. The test covered three types of reasoning: verbal, visuospatial, and numerical. The CEPS research team calculated the raw score based on the Item Response Theory and then standardized the score with a mean of zero and standard deviation of one.

Another dependent variable concerned students' academic performance. This was measured by the total scores of mid-term exams in Chinese, mathematics, and English as provided by each school. Throughout junior high school, students study these subjects, which are included in the entrance exam to senior high school. Within each cohort in one particular school, teachers adopt the same syllabus and administer the same exams during a given assessment period. Therefore, test scores for core subjects are reliable indicators of learning performance in the same school and cohort. I aligned different grading scales (full mark of 100, 120, and 150 for one subject) to the hundred-point scale. After the conversion, I checked for abnormal values outside the 0-100 range. `r printnum(num.outlier$n)` observations were outliers, accounting for less than one percent of the sample. I trimed the values of subject exam score with minimum of zero and maximum of one hundred. In this way, I can limit the impact of outliers while retaining these observations in the sample. After trimming, I aggregated the subject exam scores to obtain a total score of a 300-point scale.

@@

@@

I adopted `R` for all the analyses, the `fixest` package for econometric modeling [@berge2022], and the `modelsummary` package for presenting the results [@arel-bundock2022]. I built a two-way fixed effect model in a difference-in-differences (DID) design, following Bai and colleagues' [-@bai2018; -@bai2020] approach. The model served to analyze the educational outcome of children newly left behind by parents vis-a-vis that of children staying together with parents. I first specified an unrestricted and unadjusted model: $$\Delta score_{i,j,k} = \alpha + \beta \cdot migr_{i} + \gamma \cdot C_{j} + \lambda \cdot C_{k} + \theta \cdot score_{i,j;1} + \varepsilon_{i,j,k}$$ For student $i$ in class $j$ in baseline and class $k$ in endline, $\Delta score_{i,j}$ is the change in test score between baseline and follow-up surveys, $migr_{i}$ is the treatment dummy variable, $C_{j}$ and $C_{k}$ are respectively the classroom fixed-effects at baseline and follow-up, and $score_{i,j,k;1}$ is the test score at baseline. Classroom fixed-effect can account for differences due to time-invariant factors at the classroom level and above, as students nested in the same class are also in the same school and county. About one-sixth of the students were reassigned into new classroom at the end of baseline academic year. This necessitates adding the wave-two classroom fixed-effect.

The model is unrestricted because it does not restrict on the coefficient associated with the baseline scores. The model is unadjusted as it does not adjust for additional covariates. Theoretically, it is unnecessary to include covariates that vary over group but remain constant over time; They would cancel out in the two-way fixed effect model [@huntington-klein]. The standard errors were clustered at the classroom level.

In addition, I present an unrestricted and adjusted DID model: $$\Delta score_{i,j,k} = \alpha + \beta \cdot migr_{i} + \gamma \cdot C_{j} + \lambda \cdot C_{k} + \theta \cdot score_{i,j,k;1} + \zeta \cdot X_{i} + \varepsilon_{i,j,k}$$ where $X_{i}$ is the vector of covariates capturing the characteristics of children, their parents, and their households. The control variables were measured in the baseline survey. This version of model lifts the restriction that covariates from the baseline survey would be associated with a coefficient that equals one.

@@

The table below summarizes the changes in migration status in the sample. All `r printnum(nrow(d_cl))` children were living together with parents in one household in baseline, but about fifteen percent saw at least one parent migrated during the two waves of survey. In these new migrant households, only-mother migration and only-father migration each accounted for two out of five cases. The remaining one-fifth was both-parents migration.

```{r}
table_independent_variable
```

@@

The tables below provide the whole-sample summary statistics. In addition, the violin plots in the appendix visualize the distribution of outcome variables for the treatment group and the control group.

```{r}
table_dependent_variable
```

@@

```{r}
table_control_variable
```

@@

The unrestricted DID models support my hypothesis one -- newly left-behind adolescents are more likely to perform worse in cognitive and academic abilities than their non left-behind counterparts. If a household sees any parent migrated, the child's cognitive test score and total exam grade would decrease relative to that of children whose parents stay with them, holding everything else constant. Parental migration is linked to is `r round(coef(did_cog)[1],2)` SD in cognitive test score and `r round(coef(did_cog)[1],2)` points in academic exam grades. Both relationships are slight in strength.

@@

```{r}
table_cog
```

```{r}
table_ttl
```

@@

```{r}
table_adj_cog
```

```{r}
table_adj_ttl
```

@@

Parental migration is not a randomly assigned treatment. One technique to tackle this selection bias is matching. This method works by pairing a unit in the treatment group with one or several units in the control group based on observable covariates. Popular choices include coarsened exact matching (CEM) [@waldman2022] and nearest-neighbor propensity score matching (PSM) [@xu2015]. As @greifer2022 introduced, CEM is superior in achieving balance as it optimizes the entire joint distribution of covariates. The downside is that it leaves many units without an exact match and discards these unmatched units. The nearest-neighbor matching examines all the treated units and selects the closest control unit to be paired. The eligibility as the "closest" can be determined with the differences in propensity score. Propensity score denotes the probability of receiving treatment based on the specified covariates. Nearest-neighbor PSM does not try to optimize any condition; each pairing ignores how other pairing will or have occurred [@greifer2022]. @king2019 argued against this matching method as it risks more "imbalance, model dependence, and bias" for highly balanced data.

@@

```{r}
table_m_cog
```

```{r}
table_m_ttl
```

@@

This thesis proposed a model to assess causality between parents' migration and their stay-behind children's educational performance in the short run. I built econometric models in the difference-in-differences design to analyze data from the 2013-2014 and 2014-2015 rounds of China Education Panel Survey. Results suggest that parental migration is, in the short run, negatively associated with children's educational performance as measured by academic exams and standardized cognitive tests. I further examined whether the effect varies across different arrangements of parental migration. Results indicate that father-only migration is negatively associated with children's cognitive abilities, while mother-only migration is negatively associated with children's academic abilities. In all scenarios, the educational performance of newly left-behind children only decreased slightly compared to that of non-left-behind peers.

My research has limitations in sampling due to data constraint. Regarding sample size, the treatment group for the both-parents migration is small, lowering the statistical power. Regarding identification, I could only identify parental migration through such a condition: parental absence from a household without parental divorce or death. This is because the CEPS questionnaire asked if parents were living in the household, but did not ask why parents were absent. It is possible that some parents were jailed or hospitalized. These types of absence would be wrongly attributed to migration. Besides, I could not control for the migration history of parents, which are not included in the CEPS data. Possibly, some parents out-migrated at an earlier date, returned before the baseline survey, and migrated away again during the survey period. Their children might better cope with parental absence during the survey period, yet they would be mistaken as the newly left-behind. These limitations may bias the estimation.

@@

Let us consider the policy implications for China. The small reduction of newly left behind children's schooling performance by no means suggest that policies can remain as they are. The slight reduction is only the short-term consequence. Longer duration of parental migration may hamper educational performance to a greater extent. Besides, educational performance is only part of human development outcomes [@antia2020]. Other outcomes like mental health and non-cognitive skills are also subjected to parental absence's influence. Policymakers should reduce institutional barriers at destination cities, opening more doors for parents to migrate together with their children. Meanwhile, schools should dedicate more advisory and counseling support to children who stay behind.

@@

```{r, fig.cap = "Changes in standardized cognitive test score for newly left-behind students and non-left-behind students"}
fig_cog
```

@@

```{r, fig.cap = "Changes in academic exam score for newly left-behind students and non-left-behind students"}
fig_ttl
```

@@

```{r, fig.cap = "Changes in standardized cognitive test score by household migration arrangement"}
fig_cog_by_migration_type
```

@@

```{r, fig.cap = "Changes in academic exam score by household migration type"}
fig_ttl_by_migration_type
```

@@

```{r}
table_m_adj_cog
```

@@

```{r}
table_m_adj_ttl
```

>>>>

---
title             : "Left-Behind Children's Educational Performance Drops Slightly After Parents Migrated"
shorttitle        : "Left-Behind Children"

@@

Worldwide, millions of children are resided in their home communities but separated from their parents as either one or both parents have migrated for work. These are the so-called "left-behind children" or "stay-behind children." They are common in migrant-sending regions, such as Latin America, Sub-Sahara Africa, East Europe, and large parts of Asia. Researchers estimated that one-sixth of Mexican children [@dewaard2018], one-fifth of Bulgarian children [@popova2018], and one-fourth of Chinese children were living apart from their migrating parents [@nbs2017]. In a 2021 resolution, the Council of Europe's @parliamentaryassembly2021a expressed concern about "the scale of this phenomenon and the long-term damage it creates" (p. 2) and declared that leaving "millions of labour migrants' children without parental care is a mass violation of human rights" (p. 1).

How are stay-behind children faring when parents have migrated? Research results are mixed due to the differential effects of parental migration. Consider children's education. According to the review articles by @debrauw2019 and @vanhook2020, labor migrants send remittances back, which can fund children's schooling as well as improve their nutritional and living conditions. The inflow of remittances, however, cannot substitute for the loss of parental care. Parental migration decreases the amount and quality of parental guidance, protection, and support. It is uncertain how these two opposing effects together shape children's educational performance, which merits empirical investigation.

A challenge arises in assessing the causality between parent's migration and children's educational achievement. Experiments can unveil causal relations, but it is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Alternatively, a quasi-experimental design can help with causal inference. The cause should be on par with random assignment to participants in the analysis, conditional on identification assumptions [@cunningham2021].

In this thesis, my research question is: To what extent do different arrangements of parental migration affect the educational performance of left-behind children in the short term? I choose China as the case and analyzed data from a two-wave panel survey. The nationally representative sample includes children in both urban and rural areas. I develop a difference-in-differences model for causal inference and applied propensity score matching for robustness check. The outcome was measured with children's cognitive test score and academic exam scores. I demonstrate that parental migration reduces the educational performance of children staying behind, but only to a small extent.

@@

The other theoretical model is the family disruption model. Children flourish under parental nurturing care. This term refers to the "stable environments that promote health and adequate nutrition, protect from threats, and provide opportunities for learning and responsive, emotionally supportive and developmentally enriching relationships" [@black2021, p.1]. Parental migration decreases nurturing care, as characterized by less homework supervision and emotional support by parents [@wen2019]. Hong talked to a group of left-behind children at a rural school [@hong2019]. These students expressed the utter sense of loneliness "with no advocate, no support and with no one looking out for them" (p. 13). Many of them felt indifferent to grades and disengaged with school. Some even longed to quit school to pursue a free, independent life (p. 14).

Given the two opposing effects, researchers have obtained mixed results regarding how parental migration impacts children's education. @botezat2020 studied the case of Romania and found a positive effect of parents' migration on academic grades of adolescents aged eleven to fifteen. @wang2019 collected data from junior high school students in central China and found no effect of parental on left-behind children's achievement in mathematical test. Studying four Chinese provinces, @debrauw2017 showed that as ample migration opportunities appeared in villages, enrollment to senior high school decreased. @nguyen2016 demonstrated that parental migration was associated with lower cognitive capabilities among Indian and Vietnamese children aged five to eight. The inconsistency can be complicated by factors such as the left-behind children's age and gender, the migrant parent's original domicile and eventual destination, and the arrangement of which parent migrate for how long.

@@

Researchers have studied whether the impact differ between mother absence and father absence induced by labor migration. @xu2019 adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had "little or no association with negative outcomes" on children's academic and cognitive performances. Only-mother absence, however, showed a "strong association with negative outcomes" (p. 1646). @chen2019 conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that mother migration was "negatively associated with children's social competence and academic performance," while father migration had no direct effects on the two outcomes (pp. 860-861).

Studies noted that in some Asian and Latin American societies, the social norms designate men as breadwinners and women as caregivers. Under such norms, children are more likely to accept fathers' migration than mothers' [@murphy2022]. Left-behind children may even resent their migrant mothers for transgressing on the expected care-giving roles. I propose the second hypothesis:

$H_2$ In the short term, mother's migration negatively affect the cognitive development and academic performance of children left-behind, while father's migration does not show a effect.

@@

Three features make China a suitable case for studying left-behind children. The first feature is the huge scale of internal labor migration. According to the National Bureau of Statistics, the year 2021 recorded 292.5 million in-country migrant workers [@nbs2022]. They form the backbone of essential industries, such as retail, delivery, transportation, and manufacturing. The number of left-behind Chinese children reached 68.7 million in 2015 [@nbs2017]. The internal migration stemmed from the neoliberal economic regime.

The neoliberal economic regime sounds at odds with China, a so-called socialist country. Actually, in the 1980s, the ruling communist party ditched the planned economy, initiated market reforms, and opened up the country. Since then, China has integrated in the global economy and contributed substantial growth to it. The economic reform and globalization have enriched the coastal cities like Shanghai and Shenzhen. The rural hinterlands, however, have lagged behind. The regional inequality needs social policy to amend, but China's welfare programs are discriminatory -- it segregates the population. Those who reap the greatest benefits are public servants, urban dwellers, and formally employed people. As rural people migrate to the city, they find themselves with limited access to social welfare in the destination. The municipal governments only provide welfare to local residents and expect non-local residents access public services elsewhere [@duckett2020; @zhang2018].

The institutionalized discrimination against migrant laborers find its roots in China's "multilevel citizenship" system. As @vortherms2021 demonstrated, the constituents of citizenship -- "membership rules and rights entitlements" -- operate below the national level in China. The local state confers most of the rights and defines who is eligible to these rights through the *hukou*, or household registration system:

> A political institution rooted in centuries of Chinese bureaucracy, the modern hukou is the primary identity document defining membership and entitlements for local citizenship. Accessing fundamental citizenship rights depends on your registration status ... Registration includes basic identifying information, a specific address of registration, and a type: urban, rural, or resident. [@vortherms2021]

In other words, the hukou system binds one's citizenship rights with one's place of registry. In theory, anyone can move out of town freely. In practice, they cannot change hukou freely. When rural residents move to the city, it takes months -- if not years -- to apply for a new hukou. Before receiving urban hukou, they can hardly send their children into public schools in the host city [@gu2021]. Migrant workers had organized petitions, strikes, and protests, only to fall on deaf ears or even end with police crackdown [@chan2017]. Facing financial and policy constraints, many migrant laborers have to leave children in their hometowns.

The existing research about China's left-behind children has three major shortcomings. The first is that most literature focuses on children left behind in rural areas while neglecting the urban counterparts. The National Bureau of Statistics stated that the number of urban left-behind children had been growing and expected that the trend would continue [@nbs2017]. My sample includes urban left-behind children as well. The second shortcoming lies in sampling and research design. Some studies suffer from a limited sample size of less than one thousand participants or a limited geographic coverage of one to a few towns [@liu2021; @yang2022; @shu2021]. Others only estimated correlation using cross-sectional data [@akezhuoli2022; @jin2020]. These studies could have employed causal inference techniques like fixed effect, matching, or inverse probability of treatment weighting. The third shortcoming lies in the measurement of outcome. @chang2019 and @wang2019 measured students' educational performance with the test score of only one academic subject, namely mathematics. To tackle these limitations, I adopted a nationally representative two-wave panel dataset. I included the test results of multiple subjects to measure comprehensive academic achievement.

@@

I explored two categories of educational performance: cognitive and academic abilities. One key dependent variable, the student's cognitive skill, was measured by a fifteen-minute standardized test. The test covered three types of reasoning: verbal, visuospatial, and numerical. The CEPS research team calculated the raw score based on the Item Response Theory and then standardized the score across the whole national sample with a mean of zero and standard deviation of one [@wang2015].

Another dependent variable concerned students' academic performance. This was measured by the total scores of mid-term exams in Chinese, mathematics, and English as provided by each school. Exam scores are reliable indicators of learning performance, and are comparable across schools for the same cohort of students. This is related to how junior high schools operate as part of the compulsory education system in China. According to @oecd2016, China's junior secondary education is a unified system that does not distinguish between academic and vocational tracks. At school, Chinese, mathematics, and English are core subjects and others as supplementary subjects. Throughout junior high school, students study the core subjects in preparation for the senior high school entrance exam. Within each cohort in one particular school, teachers use the same syllabus and administer the same exams during a given assessment period. Across cities and regions, junior high schools adopt the educational content and assessment set by the Ministry of Education [@oecd2016].

I transformed the exam grades through the following steps. I first aligned different grading scales (full mark of 100, 120, and 150 for one subject) to the hundred-point scale. After the conversion, I checked for abnormal values outside the 0-100 range. `r printnum(num.outlier$n)` observations were outliers, accounting for less than one percent of the sample. I trimmed the values of subject exam score with minimum of zero and maximum of one hundred. In this way, I can limit the impact of outliers while retaining these observations in the sample. After trimming, I aggregated the subject exam scores to obtain a total score of a 300-point scale.

@@

The control variables on children include their age, gender, ethnicity, self-rated health, hukou type, number of siblings, whether the child had skipped grade or repeated grade, and the level of education that the child aspires to achieve. At the household level, I controlled the level of education of the best-educated parent, the parental expectation on the child's level of education, the perceived household financial conditions, the number of extracurricular books at home, and the internet access at home.

@@

I adopted `R` for all the analyses, the `fixest` package for econometric modeling [@berge2022], and the `modelsummary` package for presenting the results [@arel-bundock2022]. I built a two-way fixed effect model in a difference-in-differences (DID) design, following Bai and colleagues' [-@bai2018; -@bai2020] approach. The model served to analyze the educational outcome of children newly left behind by parents vis-a-vis that of children staying together with parents. I first specified an unrestricted and unadjusted model: $$\Delta score_{i,j,k} = \alpha + \beta \cdot migr_{i} + \gamma \cdot score_{i,j;1} + \lambda \cdot C_{j} + \theta \cdot C_{k} + \varepsilon_{i,j,k}$$ For student $i$ in class $j$ in baseline and class $k$ in endline, $\Delta score_{i,j}$ is the change in test score between baseline and follow-up surveys, $migr_{i}$ is the treatment dummy variable, $score_{i,j,k;1}$ is the test score at baseline, and $C_{j}$ and $C_{k}$ are the classroom fixed-effects at baseline and follow-up, respectively. Classroom fixed-effect can account for differences due to time-invariant factors at the classroom level and above, as students nested in the same class are also in the same school and county. About one-sixth of the students were reassigned into new classroom at the end of baseline academic year. This necessitates adding the wave-two classroom fixed-effect.

The model is unrestricted because it does not restrict on the coefficient associated with the baseline scores. The model is unadjusted as it does not adjust for additional covariates. Theoretically, it is unnecessary to include covariates that vary over group but remain constant over time; They would cancel out in the two-way fixed effect model [@huntington-klein2022]. The standard errors were clustered at the classroom level.

In addition, I present an unrestricted and adjusted DID model: $$\Delta score_{i,j,k} = \alpha + \beta \cdot migr_{i} + \gamma \cdot score_{i,j,k;1} + \lambda \cdot C_{j} + \theta \cdot C_{k} + \zeta \cdot X_{i} + \varepsilon_{i,j,k}$$ where $X_{i}$ is the vector of covariates capturing the characteristics of children, their parents, and their households. The control variables were measured in the baseline survey. This version of model lifts the restriction that covariates from the baseline survey would be associated with a coefficient that equals one.

@@

Table 1 below summarizes the changes in migration status in the sample. All `r printnum(nrow(d_cl))` children were living together with parents in one household in baseline, but about fifteen percent saw at least one parent migrated during the two waves of survey. In these new migrant households, only-mother migration and only-father migration each accounted for two out of five cases. The remaining one-fifth was both-parents migration.

```{r i-var}
table_independent_variable
```

@@

Tables 2 and 3 below provide the whole-sample summary statistics. In addition, the violin plots in the appendix visualize the distribution of outcome variables for the treatment group and the control group.

```{r d-var}
table_dependent_variable
```

@@

```{r c-var}
table_control_variable
```

@@

The unrestricted DID models support my hypothesis one -- newly left-behind adolescents are more likely to perform worse in cognitive and academic abilities than their non left-behind counterparts. If a household sees any parent migrated, the child's cognitive test score and total exam grade would decrease relative to that of children whose parents stay with them, holding everything else constant. Parental migration is linked to `r round(coef(did_cog)[1],2)` SD in cognitive test score and `r round(coef(did_cog)[1],2)` points in academic exam grades. Both relationships are slight in strength.

@@

```{r reg-cog}
table_cog
```

```{r reg-ttl}
table_ttl
```

@@

```{r reg-cog-adj}
table_adj_cog
```

```{r reg-ttl-adj}
table_adj_ttl
```

@@

Parental migration is not a randomly assigned treatment. A possible circumstance is that smart and ambitious students perform better at school; their high performance and ambition prompt parents to migrate to support children's schooling. Another scenario is that poverty may hamper children's educational achievement while driving parents to migrate. To tackle the selection bias herein, an applicable technique is matching. This method makes two groups comparable by pairing a unit in the treatment group with one or several units in the control group based on observable covariates.

Popular choices for matching include coarsened exact matching (CEM) [@waldman2022] and nearest-neighbor propensity score matching (PSM) [@xu2015]. As @greifer2022 introduced, CEM is superior in achieving balance as it optimizes the entire joint distribution of covariates. The downside is that it leaves many units without an exact match and discards these unmatched units. The nearest-neighbor matching examines all the treated units and selects the closest control unit to be paired. The eligibility as the "closest" can be determined with the differences in propensity score. Propensity score denotes the probability of receiving treatment based on the specified covariates. Nearest-neighbor PSM does not try to optimize any condition; each pairing ignores how other pairing will occur or have occurred [@greifer2022]. @king2019 argued against this matching method as it risks more "imbalance, model dependence, and bias" for highly balanced data.

@@

```{r reg-m-cog}
table_m_cog
```

```{r reg-m-ttl}
table_m_ttl
```

@@

This thesis proposes a model to infer causality between parents' migration and their stay-behind children's educational performance in the short run. I build econometric models in the difference-in-differences design to analyze data from the 2013-2014 and 2014-2015 rounds of China Education Panel Survey. Results suggest that parental migration is, in the short run, negatively associated with children's educational performance as measured by academic exams and standardized cognitive tests. I further examine whether the effect varies across different arrangements of parental migration. Results indicate that father-only migration is negatively associated with children's cognitive abilities, while mother-only migration is negatively associated with children's academic abilities. In all scenarios, the educational performance of newly left-behind children only decreases slightly compared to that of non-left-behind peers.

My research has limitations in sampling due to data constraint. Regarding sample size, the treatment group for the both-parents migration is small (n = 131), making the statistical power low. Regarding identification, I could only identify parental migration through such a condition: parental absence from a household without parental divorce or death. This is because the CEPS questionnaire asked if parents were living in the household, but did not ask why parents were absent. If some parents were jailed or hospitalized, their absence would be wrongly attributed to migration. Besides, I could not control for the migration history of parents, which are not included in the CEPS data. Possibly, some parents out-migrated at an earlier date, returned before the baseline survey, and migrated away again during the survey period. Their children might better cope with parental absence during the survey period, yet they would be mistaken as the newly left-behind. These limitations may bias the estimation.

@@

Let us consider the policy implications for China. The government has pushed for reforming the multilevel citizenship scheme. By 2022, small- and medium-sized cities have liberalized the application for local, urban hukou. However, the most popular migration destinations are metropolises like Beijing and Shanghai. These mega-cities still restrict migrant workers' local citizenship rights [@wang2020]. The hukou scheme shall not remain as it is, despite the small reduction of newly left behind children's schooling performance. The slight reduction is only the short-term consequence. Longer duration of parental migration may hamper educational performance to a greater extent. Besides, educational attainment is only part of human development outcomes. Other outcomes like mental health and non-cognitive skills are also subjected to parental absence's influence [@antia2020]. Policymakers should accelerate the reforms on the multilevel citizenship scheme, opening more doors for parents to migrate together with their children. Meanwhile, schools should dedicate more advisory and counseling support to children who stay behind.

@@

(ref:cog-dif) Changes in standardized cognitive test score

(ref:ttl-dif) Changes in academic exam score

(ref:migr) for new left-behind students and non-left-behind students

(ref:migr-type) by household migration arrangement

```{r v-plot-cog, fig.cap = paste("(ref:cog-dif)", "(ref:migr)")}
fig_cog
```

@@

```{r v-plot-ttl, fig.cap = paste("(ref:ttl-dif)", "(ref:migr)")}
fig_ttl
```

@@

```{r v-plot-cog-1, fig.cap = paste("(ref:cog-dif)", "(ref:migr-type)")}
fig_cog_by_migration_type
```

@@

```{r v-plot-ttl-1, fig.cap = paste("(ref:cog-dif)", "(ref:migr-type)")}
fig_ttl_by_migration_type
```

@@

(ref:cobalt-match) Covariate balance before and after matching

(ref:cobalt-any) by any parent migration

(ref:cobalt-fa) by father-only migration

(ref:cobalt-mo) by mother-only migration

(ref:cobalt-both) by both parents migration

```{r cobal-any, fig.cap = paste("(ref:cobalt-match)", "(ref:cobalt-any)")}
love.plot(
  m.out,
  threshold = c(m = .1),
  binary = "std", 
  limits = c(0, 0.55),
  abs = T
  )
```

@@

```{r cobal-mo, fig.cap = paste("(ref:cobalt-match)", "(ref:cobalt-mo)")}
love.plot(
  m.out_mo_only,
  threshold = c(m = .1),
  binary = "std", 
  limits = c(0, 0.55),
  abs = T
  )
```

```{r cobal-fa, fig.cap = paste("(ref:cobalt-match)", "(ref:cobalt-fa)")}
love.plot(
  m.out_fa_only,
  threshold = c(m = .1),
  binary = "std", 
  limits = c(0, 0.55),
  abs = T
  )
```

```{r cobal-both, fig.cap = paste("(ref:cobalt-match)", "(ref:cobalt-both)")}
love.plot(
  m.out_both,
  threshold = c(m = .1),
  binary = "std", 
  limits = c(0, 0.55),
  abs = T
  )
```

```{r reg-m-adj-cog}
table_m_adj_cog
```

```{r reg-m-adj-ttl}
table_m_adj_ttl
```
